Subset coercions in \Coq

### 2006-05-02 papers/article/main.tex
One minute fixes

\subsection{Typing and conversion}
Our adaptation of \ps{} differs from the presentation of
\ps{} in \cite{Rushby98:TSE, Shankar&Owre:WADT99} essentially because of the
proof terms, but we have also extended it in some ways.

@@

\Russell{}'s typing system (figure \vref{fig:typing-decl-rules}) is based
on the Calculus of Constructions \cite{CoquandHuet,Barras99}, with sigma types but without universes.
This restriction may be removed in future work, but cause no problem for
our main purpose which is programming. The judgement $`G \typed t : T$
means $t$ is a well-typed term of type $T$ in environment $`G$, and $`G
\typed T \subd U$ means $T$ is convertible to $U$ in environment $`G$. 
Following the presentation as a pure type system \cite{PTS}, 
the set of sorts $\PTSsorts$ is defined as $\setproptype$. As usual, we
let $s, s_i$ for $i `: `N$ range over sorts. The
axioms are $\PTSaxioms = \{ (\Set, \Type), (\Prop, \Type) \}$ and the
rules $\PTSrules$ are defined by the following (functional) relation:
\begin{table}[h]
  \vspace{-1.5em}
  \axiomsd
  \vspace{-2em}
  \caption{Definition of $\PTSrules$}
  \label{fig:PTSrules}
\end{table}

@@

\subsubsection{Equivalence}
We have renamed the technique from \ps{} to 
subset \emph{equivalence} because we have a symmetric relation, contrary
to usual subtyping relations. It also permits to include $\beta$-conversion
directly in the relation. The judgment $`G \typed T \subd{} U$ (figure
\vref{fig:conversion-decl-rules}) reads $T$ is convertible to $U$ in
environment $`G$. In the following, it is always assumed that the two
input types of the conversion are well-sorted, but maybe not by the same sort.

@@

The rule \irule{SubConv} integrates $\beta\pi$-conversion in the
judgment. It is well-known \cite{Luo, Pottinger?} that $\beta\pi$ is
confluent only on well typed terms, so we have to ensure that our rules,
when read from bottom to top, keep well-typedness. That's the purpose of
the $`G \typed U : s$ premise of the \irule{SubTrans} rule. This rules
ensures that our judgement constructs an \emph{equivalence} and has
proper modularity. The next two rules do context closure for dependent
products and sums. It is remarkable that we use contravariance for
domains in the \irule{SubProd} rule, not restricting to invariance as in
\PVS \cite{journals/toplas/Castagna95}. It is accessory here, as we could have used \emph{co}variance and
still get the same judgements because we have symetry. However, it will
become important when we create coercions (see
\vref{fig:subtyping-impl-rules}).

The really interesting rules are \irule{SubSub} and
\irule{SubProof}. The first one allows to use an object of a subset type
as an object of its support type. The latter allows (maybe abusively) to
consider an object of any type as an object of any subset based on
this type. By the assumption on the input types of the judgment, we know
that the property $P$ is well-formed, but we do not care about the proof.

>>>>

\subsection{Typing and conversion}
Our adaptation of \ps{} differs from the presentation of
given in \cite{Rushby98:TSE, Shankar&Owre:WADT99} essentially because of the
proof terms, but we have also extended it in some ways.

@@

\Russell{}'s typing system (figure \vref{fig:typing-decl-rules}) is based
on the Calculus of Constructions \cite{CoquandHuet,Barras99}, with sigma types but without universes.
This restriction may be removed in future work, but cause no problem for
our main purpose which is programming. The judgement $`G \typed t : T$
means $t$ is a well-typed term of type $T$ in environment $`G$, and $`G
\typed T \subd U$ means $T$ is equivalent to $U$ in environment $`G$. 
Following the presentation as a pure type system \cite{PTS}, 
the set of sorts $\PTSsorts$ is defined as $\setproptype$. As usual, we
let $s, s_i$ for $i `: `N$ range over sorts. The
axioms are $\PTSaxioms = \{ (\Set, \Type), (\Prop, \Type) \}$ and the
set of rules $\PTSrules$ is defined by the following (functional) relation:
\begin{table}[h]
  \vspace{-1.5em}
  \axiomsd
  \vspace{-2em}
  \caption{Definition of $\PTSrules$}
  \label{fig:PTSrules}
\end{table}

@@

\subsubsection{Equivalence}
We have renamed the technique from \ps{} to 
\emph{subset equivalence} because we have a symmetric relation, contrary
to usual subtyping relations. It also permits to include $\beta$-conversion
directly in the relation. The judgment $`G \typed T \subd{} U$ (figure
\vref{fig:conversion-decl-rules}) reads $T$ is convertible to $U$ in
environment $`G$. In the following, it is always assumed that the two
input types of the equivalence judgment are well-sorted, but maybe not by the same sort.

@@

The rule \irule{SubConv} integrates $\beta\pi$-conversion in the
judgment. It is well-known \cite{Luo, Pottinger?} that $\beta\pi$ is
confluent and terminating only on well typed terms, so we have to
ensure that our rules, when read from bottom to top, keep
well-typedness. That's the purpose of
the $`G \typed U : s$ premise of the \irule{SubTrans} rule. This rule
ensures that our judgement constructs an \emph{equivalence} and has
proper modularity. The next two rules do context closure for dependent
products and sums. It is remarkable that we use contravariance
\cite{journals/toplas/Castagna95} for
domains in the \irule{SubProd} rule, not restricting to invariance as in
\PVS. It is accessory here, as we could have used \emph{co}variance and
still get the same judgements because we have symetry. However, it will
become important when we create coercions (see
\vref{fig:subtyping-impl-rules}).

The really interesting rules are \irule{SubSub} and
\irule{SubProof}. The first one allows to use an object of a subset type
as an object of its support type. The later allows (maybe abusively) to
consider an object of any type as an object of any subset based on
this type. By the assumption on the input types of the judgment, we know
that the property $P$ is well-formed, but we do not care about the proof.
